---
title: "Pre-processing"
output: html_notebook
Note: This notebook contains a manipulation of the 3 datasets, leads, contacts, and marketing.
---

# Intial manipulations

## Libraries 
```{r}
library(tidyr)
library(dplyr)
library("readr")
```

## Read CSV's
```{r}
tudor_leads <- read_csv("Extracted data Aaron/tudor_leads.csv")
tudor_export_contacts <- read_csv("Extracted data Aaron/tudor_export_contacts.csv")
tudor_marketing_events <- read_csv("Extracted data Aaron/tudor_marketing_events.csv")
```

## Rename datasets
```{r}
leads <- tudor_leads
contacts <- tudor_export_contacts
marketing <- tudor_marketing_events
```

## De-duplicate contacts dataframe
```{r}
# Proof that it contains duplicates and visualization of dataframe with these
duplicates <- contacts[duplicated(contacts$contact_id),]

# De-duplication
contacts <- contacts %>% distinct(contact_id, .keep_all = TRUE)
```

## Pre-process column opportunityid from contacts
```{r}
contacts$opportunityid <- ifelse(is.na(contacts$opportunityid), 0, 1)
```

## Summary stats on opoortunityid
```{r}
table(contacts$opportunityid)
17033/ (17033+82928)
17033+82928
```

# Marketing

## Manipulating marketing_events database

### Unique combinations marketing events
```{r}
# Use the unique() function to get distinct combinations of activity_type and new_status
distinct_combinations <- unique(marketing[c("activity_type", "new_status")])
# Print the results
print(distinct_combinations)
```

### Adding up marketing events
```{r}
# here I am dding up marketing events, this is the mistake that I made, since I got 0 or 1 only for marketing events. Actually, I should have done the same that Aaaron did with platform events
# but first, I need to deduplicate it, based on the 4 columns date, lead id and the two unique identifiers
marketing_subset <- marketing[, c(2, 3, 5, 6)]

# Count the number of duplicate rows in the marketing data set
n_duplicates <- sum(duplicated(marketing_subset))
cat("The marketing data set has", n_duplicates, "duplicate rows.\n")

# Remove duplicates from the subsetted data frame
marketing_subset_unique <- marketing_subset[!duplicated(marketing_subset), ]

# now I can add up the events, and get rid of time column
# Aggregate the data
aggregated_data <- marketing_subset_unique %>%
  group_by(leadorcontactid, activity_type, new_status) %>%
  summarise(num_occurrences = n()) %>%
  ungroup()

# View the aggregated data
print(aggregated_data)

# Rename dataframe to marketing_subset_unique so that I don't have to change everything again
marketing_subset_unique <- aggregated_data
```

### Now we do the pivot, when there are no more duplicates
```{r}
wide_data <- pivot_wider(data = marketing_subset_unique, 
                         id_cols = leadorcontactid, 
                         names_from = c(activity_type, new_status), 
                         values_from = num_occurrences, 
                         #values_fill = list(flag = as.list(integer(1))), 
                         names_sep = "_")

# Replace missing values in columns 2 to 51 of wide_table with 0's
wide_data[, 2:51][is.na(wide_data[, 2:51])] <- 0
```

### Merging wide marketing dataset with the leads and contacts tables and summary stats
```{r}
# -- merging with leads

# Merge the two data frames using a left join on the "lead_id" column
marketing_merged_leads <- merge(leads, wide_data, by.x = "lead_id", by.y = "leadorcontactid", all.x = TRUE)

# Count the number of rows with NAs in columns 11-60 of marketing_merged_leads
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 11:60])) > 0)
cat("There are", na_count, "observations with NAs in columns 11-60 of marketing_merged_leads.\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(marketing_merged_leads))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 2:60])) > 0)
na_count

# Percentage of completion taking in consideration accounts and platform_events merging
cat("precentage completion leads:", (1-na_count/nrow(marketing_merged_leads))*100)

# -- merging with contacts

# Merge the two data frames using a left join on the "contact_id" column
marketing_merged_contacts <- merge(contacts, wide_data, by.x = "contact_id", by.y = "leadorcontactid", all.x = TRUE)

# Count the number of rows with NAs in columns 11-60 of marketing_merged_contacts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 12:61])) > 0)
cat("There are", na_count, "observations with NAs in columns 12-61 of marketing_merged_contacts.\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(marketing_merged_contacts))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 2:61])) > 0)
na_count

# Percentage of completion taking in consideration accounts and marketing_events merging
cat("precentage completion contacts:", (1-na_count/nrow(marketing_merged_contacts))*100)
```

## Finalizing

### Finally subsetting for NA's for both leads and contacts
```{r}
# -- leads
subset_marketing_merged_leads <- marketing_merged_leads[complete.cases(marketing_merged_leads),]

# -- contacts
subset_marketing_merged_contacts <- marketing_merged_contacts[complete.cases(marketing_merged_contacts),]
```

### Analyzing opportunityid
```{r}
#-> count how many opportunityid have 1 and 0
table(subset_marketing_merged_contacts$opportunityid)
```

### Merging leads and contacts tables
```{r}
# First create opportunityid column for lead table
subset_marketing_merged_leads <- subset_marketing_merged_leads %>%
  mutate(opportunityid = 0) %>% # create new column opportunityid with value 0
  select(1:10, opportunityid, 11:ncol(subset_marketing_merged_leads)) # move the new column to the 11th position

# Merge leads and contacts

# First rename first column from each dataset
subset_marketing_merged_leads <- subset_marketing_merged_leads %>%
  rename(id = lead_id)

subset_marketing_merged_contacts <- subset_marketing_merged_contacts %>%
  rename(id = contact_id)

# check if column names are equal
all(names(subset_marketing_merged_leads) == names(subset_marketing_merged_contacts))

# Concatenate the two
if(all(names(subset_marketing_merged_leads) == names(subset_marketing_merged_contacts))) {
  combined_df <- rbind(subset_marketing_merged_leads, subset_marketing_merged_contacts)
} else {
  # if the column names and order are not the same, show an error message
  stop("Column names or order are not the same for both data frames.")
}
```

## Export dataframe only containing marketing
```{r}
write.csv(combined_df, file = "dataset_marketing.csv", row.names = FALSE)
```