---
title: "Adding marketing_events and platform_events to leads and contacts"
output: html_notebook
Note: This notebook contains a manipulation of the 4 datasets, leads, contacts, marketing and platform. Now, because the decision was taken to only include the marketing dataset, I can directly export the dataset_marketing2 to Python.
---

# Intial manipulations

## Libraries 
```{r}
library(tidyr)
library(dplyr)
library("readr")
```

## Read CSV's
```{r}
tudor_leads <- read_csv("Extracted data Aaron/tudor_leads.csv")
tudor_export_contacts <- read_csv("Extracted data Aaron/tudor_export_contacts.csv")
tudor_marketing_events <- read_csv("Extracted data Aaron/tudor_marketing_events.csv")
tudor_platform_events <- read_csv("Extracted data Aaron/tudor_platform_events.csv")
```

## Rename datasets
```{r}
leads <- tudor_leads
contacts <- tudor_export_contacts
marketing <- tudor_marketing_events
platform <- tudor_platform_events
```

## De-duplicate contacts dataframe
```{r}
# Proof that it contains duplicates and visualization of dataframe with these
duplicates <- contacts[duplicated(contacts$contact_id),]

# De-duplication
contacts <- contacts %>% distinct(contact_id, .keep_all = TRUE)
```

## Pre-process column opportunityid from contacts
```{r}
contacts$opportunityid <- ifelse(is.na(contacts$opportunityid), 0, 1)
```

## Summary stats on opoortunityid
```{r}
table(contacts$opportunityid)
17033/ (17033+82928)
17033+82928
```

# Marketing

## Manipulating marketing_events database

### Unique combinations marketing events
```{r}
# Use the unique() function to get distinct combinations of activity_type and new_status
distinct_combinations <- unique(marketing[c("activity_type", "new_status")])
# Print the results
print(distinct_combinations)
```

### Adding up marketing events
```{r}
# here I am dding up marketing events, this is the mistake that I made, since I got 0 or 1 only for marketing events. Actually, I should have done the same that Aaaron did with platform events
# but first, I need to deduplicate it, based on the 4 columns date, lead id and the two unique identifiers
marketing_subset <- marketing[, c(2, 3, 5, 6)]

# Count the number of duplicate rows in the marketing data set
n_duplicates <- sum(duplicated(marketing_subset))
cat("The marketing data set has", n_duplicates, "duplicate rows.\n")

# Remove duplicates from the subsetted data frame
marketing_subset_unique <- marketing_subset[!duplicated(marketing_subset), ]

# now I can add up the events, and get rid of time column
# Aggregate the data
aggregated_data <- marketing_subset_unique %>%
  group_by(leadorcontactid, activity_type, new_status) %>%
  summarise(num_occurrences = n()) %>%
  ungroup()

# View the aggregated data
print(aggregated_data)

# Rename dataframe to marketing_subset_unique so that I don't have to change everything again
marketing_subset_unique <- aggregated_data
```

### Now we do the pivot, when there are no more duplicates
```{r}
wide_data <- pivot_wider(data = marketing_subset_unique, 
                         id_cols = leadorcontactid, 
                         names_from = c(activity_type, new_status), 
                         values_from = num_occurrences, 
                         #values_fill = list(flag = as.list(integer(1))), 
                         names_sep = "_")

# Replace missing values in columns 2 to 51 of wide_table with 0's
wide_data[, 2:51][is.na(wide_data[, 2:51])] <- 0
```

### Merging wide marketing dataset with the leads and contacts tables and summary stats
```{r}
# -- merging with leads

# Merge the two data frames using a left join on the "lead_id" column
marketing_merged_leads <- merge(leads, wide_data, by.x = "lead_id", by.y = "leadorcontactid", all.x = TRUE)

# Count the number of rows with NAs in columns 11-60 of marketing_merged_leads
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 11:60])) > 0)
cat("There are", na_count, "observations with NAs in columns 11-60 of marketing_merged_leads.\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(marketing_merged_leads))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(marketing_merged_leads[, 2:60])) > 0)
na_count

# Percentage of completion taking in consideration accounts and platform_events merging
cat("precentage completion leads:", (1-na_count/nrow(marketing_merged_leads))*100)

# -- merging with contacts

# Merge the two data frames using a left join on the "contact_id" column
marketing_merged_contacts <- merge(contacts, wide_data, by.x = "contact_id", by.y = "leadorcontactid", all.x = TRUE)

# Count the number of rows with NAs in columns 11-60 of marketing_merged_contacts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 12:61])) > 0)
cat("There are", na_count, "observations with NAs in columns 12-61 of marketing_merged_contacts.\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(marketing_merged_contacts))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(marketing_merged_contacts[, 2:61])) > 0)
na_count

# Percentage of completion taking in consideration accounts and marketing_events merging
cat("precentage completion contacts:", (1-na_count/nrow(marketing_merged_contacts))*100)
```

## Finalizing

### Finally subsetting for NA's for both leads and contacts
```{r}
# -- leads
subset_marketing_merged_leads <- marketing_merged_leads[complete.cases(marketing_merged_leads),]

# -- contacts
subset_marketing_merged_contacts <- marketing_merged_contacts[complete.cases(marketing_merged_contacts),]
```

### Analyzing opportunityid
```{r}
#-> count how many opportunityid have 1 and 0
table(subset_marketing_merged_contacts$opportunityid)
```

### Merging leads and contacts tables
```{r}
# First create opportunityid column for lead table
subset_marketing_merged_leads <- subset_marketing_merged_leads %>%
  mutate(opportunityid = 0) %>% # create new column opportunityid with value 0
  select(1:10, opportunityid, 11:ncol(subset_marketing_merged_leads)) # move the new column to the 11th position

# Merge leads and contacts

# First rename first column from each dataset
subset_marketing_merged_leads <- subset_marketing_merged_leads %>%
  rename(id = lead_id)

subset_marketing_merged_contacts <- subset_marketing_merged_contacts %>%
  rename(id = contact_id)

# check if column names are equal
all(names(subset_marketing_merged_leads) == names(subset_marketing_merged_contacts))

# Concatenate the two
if(all(names(subset_marketing_merged_leads) == names(subset_marketing_merged_contacts))) {
  combined_df <- rbind(subset_marketing_merged_leads, subset_marketing_merged_contacts)
} else {
  # if the column names and order are not the same, show an error message
  stop("Column names or order are not the same for both data frames.")
}
```

## Export dataframe only containing marketing
```{r}
write.csv(combined_df, file = "dataset_marketing.csv", row.names = FALSE)
```


# Platform

## Subset and delete observations which have NA for id
```{r}
platform <- platform [, c(1, 3, 4)]
# count the number of observations with missing salesforce_id
n_missing <- sum(is.na(platform$salesforce_id))

# subset the dataset to only include non-missing salesforce_id
platform_new <- platform[!is.na(platform$salesforce_id),]

# count the number of observations that were removed
n_deleted <- nrow(platform) - nrow(platform_new)

# print the number of observations removed
cat("Number of observations removed: ", n_deleted, "\n")

# print the number of remaining observations
cat("Number of remaining observations: ", nrow(platform_new), "\n")

# Renaming new deleted NA's dataset back to platform
platform <- platform_new
```

## Manipulating platform database

### All unique platform events 
```{r}
# Use the unique() function to get distinct combinations of activity_type and new_status
distinct_combinations1 <- unique(platform[c("event_type")])
# Print the results
print(distinct_combinations1)

# write.csv(distinct_combinations1, file = "platform_events.csv", row.names = FALSE) # for Manon
```

### Checking for duplicates
```{r}
platform_subset <- platform [,c(1,2)]
# Count the number of duplicate rows in the marketing data set
n_duplicates <- sum(duplicated(platform_subset))
# Print the result
cat("The platform data set has", n_duplicates, "duplicate rows.\n")
```

### Aggregating dataset 
```{r}
platform_agg <- aggregate(events ~ salesforce_id + event_type, data = platform, sum)
platform <- platform_agg
```

### Pivot the database
```{r}
platform_wide <- platform %>% 
  pivot_wider(names_from = event_type, values_from = events, id_cols = salesforce_id)

# Replace missing values in columns 2 to 273 of wide_table with 0's
platform_wide[, 2:273][is.na(platform_wide[, 2:273])] <- 0
```

## Merging wide platform dataset with the leads and contacts tables
```{r}
# -- merging with leads

# Merge the two data frames using a left join on the "lead_id" column
platform_merged_leads <- merge(leads, platform_wide, by.x = "lead_id", by.y = "salesforce_id", all.x = TRUE)

# Count the number of rows with NAs in columns 11-282 of merged_leads
na_count <- sum(rowSums(is.na(platform_merged_leads[, 11:282])) > 0)
cat("There are", na_count, "observations with NAs in columns 11-282 of platform_merged_leads\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(platform_merged_leads[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(platform_merged_leads))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(platform_merged_leads[, 2:282])) > 0)
na_count

# Percentage of completion taking in consideration accounts and platform_events merging
cat("precentage completion leads:", (1-na_count/nrow(platform_merged_leads))*100)

# -- merging with contacts

# Merge the two data frames using a left join on the "lead_id" column
platform_merged_contacts <- merge(contacts, platform_wide, by.x = "contact_id", by.y = "salesforce_id", all.x = TRUE)

# Count the number of rows with NAs in columns 11-282 of merged_leads
na_count <- sum(rowSums(is.na(platform_merged_contacts[, 12:283])) > 0)
cat("There are", na_count, "observations with NAs in columns 11-282 of platform_merged_contacts\n")

# Count the number of rows with NAs that come due to accounts
na_count <- sum(rowSums(is.na(platform_merged_contacts[, 2:9])) > 0)
cat("There are", na_count, "observations with NAs due to accounts out of", nrow(platform_merged_contacts))

# Count the number of rows with NAs that come due to both platform and accounts
na_count <- sum(rowSums(is.na(platform_merged_contacts[, 2:283])) > 0)
na_count

# Percentage of completion taking in consideration accounts and platform_events merging
cat("precentage completion contacts:", (1-na_count/nrow(platform_merged_contacts))*100)
```

## Finalizing

### Finally subsetting for NA's for both leads and contacts
```{r}
subset_platform_merged_leads <- platform_merged_leads[complete.cases(platform_merged_leads),]

subset_platform_merged_contacts <- platform_merged_contacts[complete.cases(platform_merged_contacts),]
```


### Analyzing opportunityid
```{r}
#-> count how many opportunityid have 1 and 0
table(subset_platform_merged_contacts$opportunityid)
4881/(2632+4881)
```

### Merging leads and contacts tables
```{r}
# First create opportunityid column for lead table
subset_platform_merged_leads <- subset_platform_merged_leads %>%
  mutate(opportunityid = 0) %>% # create new column opportunityid with value 0
  select(1:10, opportunityid, 11:ncol(subset_platform_merged_leads)) # move the new column to the 11th position

# Merge leads and contacts

# First rename first column from each dataset
subset_platform_merged_leads <- subset_platform_merged_leads %>%
  rename(id = lead_id)

subset_platform_merged_contacts <- subset_platform_merged_contacts %>%
  rename(id = contact_id)

# check if column names are equal
all(names(subset_platform_merged_leads) == names(subset_platform_merged_contacts))

# Concatenate the two
if(all(names(subset_platform_merged_leads) == names(subset_platform_merged_contacts))) {
  combined_df_platform <- rbind(subset_platform_merged_leads, subset_platform_merged_contacts)
} else {
  # if the column names and order are not the same, show an error message
  stop("Column names or order are not the same for both data frames.")
}
```


# Export dataframe containing both marketing and platform

## Combining the two
```{r}
# First checking for duplicate, because apprently it contains quite a few
has_duplicates <- any(duplicated(combined_df$id))
duplicate_rows <- combined_df[duplicated(combined_df$id),]

# Count the number of duplicate rows
num_duplicates <- sum(duplicated(combined_df$id))

# Renaming combine_df's
marketing_combined <- combined_df
platform_combined <- combined_df_platform

# Counting common rows (result: 52155)
common_ids <- intersect(marketing_combined$id, platform_combined$id)
num_common_ids <- length(common_ids)

# Opp ID count
num_opportunity_1 <- sum(marketing_combined$opportunityid == 1)
num_opportunity_2 <- sum(platform_combined$opportunityid == 1)

# Joining
# get the columns that are unique to each data frame
marketing_cols <- setdiff(names(marketing_combined), names(platform_combined))
platform_cols <- setdiff(names(platform_combined), names(marketing_combined))
# perform inner join and rename columns
common_df <- inner_join(marketing_combined, platform_combined, by = "id") %>%
  rename_with(~paste0("M_", .), intersect(names(.), marketing_cols)) %>%
  rename_with(~paste0("P_", .), intersect(names(.), platform_cols))

# Counting how many have opportunityid = 1 in common_df (meaning the dataframe that has leads, contacts and all the variables from marketing and platform)
num_rows_with_oppid_1 <- common_df %>%
  filter(opportunityid.x == 1) %>%
  nrow()

num_rows_with_oppid_1 <- marketing_combined %>%
  filter(opportunityid == 1) %>%
  nrow()
```

## Rename columns
```{r}
# Renaming columns that are the same for both marketing and pltaform so that they don't have "_x" termination
colnames(common_df)[colnames(common_df) == "accountid.x"] <- "accountid"
colnames(common_df)[colnames(common_df) == "account_sales_unit__c.x"] <- "account_sales_unit__c"
colnames(common_df)[colnames(common_df) == "gtm_segment__c.x"] <- "gtm_segment__c"
colnames(common_df)[colnames(common_df) == "gtm_industry__c.x"] <- "gtm_industry__c"
colnames(common_df)[colnames(common_df) == "region__c.x"] <- "region__c"
colnames(common_df)[colnames(common_df) == "sales_unit_group__c.x"] <- "sales_unit_group__c"
colnames(common_df)[colnames(common_df) == "name.x"] <- "name"
colnames(common_df)[colnames(common_df) == "country_grouping__c.x"] <- "country_grouping__c"
colnames(common_df)[colnames(common_df) == "createddate.x"] <- "createddate"
colnames(common_df)[colnames(common_df) == "opportunityid.x"] <- "opportunityid"
```

## Removing duplicate columns
```{r}
# Position of the duplicate columns from the platform dataset that we have to remove
which(names(common_df) == "accountid.y")
# Removing them
common_df <- common_df[, -c(62:71)]
```

## Export
```{r}
write.csv(common_df, file = "Dataframe marketing and platform.csv", row.names = FALSE)
```

